---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Prior to the analysis, we must first load the data and clean it.

```{r message = FALSE}
# Load the necessary packages
library(tidyverse)
library(magrittr)
library(zeallot)

# Load the data
activity_data <- read_csv('activity_data.csv')

# Glimpse the unprocessed data
activity_data %>% glimpse()
```

Notice that the `date` column has already been converted into a `Datetime`
object by the `read_csv` function. It's quite clear from the above data frame
summary that there are `NA`s in the `step` column. We will deal with them by
simply converting them into zeros.

```{r}
# Convert all NAs in the 'steps' column to 0
activity_data %<>% replace_na(replace = list(steps = 0))

# Glimpse the data to verify that the conversion worked
activity_data %>% glimpse()
```


## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
